Illusion Game & PCS

### 2025-04-30 analysis/1_cleaning.qmd
updated analysis script

ipip6 <- select(data_ppt, starts_with("ipip_"))
ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]

@@

pid5 <- select(data_ppt, starts_with("pid_"))

@@

```{r p_ethnicity}
ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
  ggplot(aes(x, y, fill = group)) + 
  ggwaffle::geom_waffle() + 
  coord_equal() + 
  scale_fill_flat_d() + 
  ggwaffle::theme_waffle() +
  labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2)),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

>>>>

```{r}
ipip6 <- select(data_ppt, starts_with("ipip_"))
ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]

@@

```{r}
pid5 <- select(data_ppt, starts_with("pid_"))

@@

```{r p_ethnicity}
# ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
#   ggplot(aes(x, y, fill = group)) + 
#   ggwaffle::geom_waffle() + 
#   coord_equal() + 
#   scale_fill_flat_d() + 
#   ggwaffle::theme_waffle() +
#   labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
#   theme(
#     plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
#     plot.subtitle = element_text(size = rel(1.2)),
#     axis.title.x = element_blank(),
#     axis.title.y = element_blank()
#   )
```

### 2025-04-30 analysis/2_analysis.qmd
updated analysis script

@@

@@

**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended

@@

  return(list(
    ERcongruent_vs_mild = ERcongruent_vs_mild,
    ERmild_vs_strong = ERmild_vs_strong,
    IEScongruent_vs_mild = IEScongruent_vs_mild,
    IESmild_vs_strong = IESmild_vs_strong
  ))
}

````

@@

ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
ttest <- compute_ttest(data_vs, "MullerLyer")
ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
```

@@

er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)

er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
bf_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)

er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
er_verticalhorizontal_post_ms <- describe_posterior(ttestcor_verticalhorizontal$ERmild_vs_strong)
```

@@

IES_ebbinghaus_post_cm <- describe_posterior(cor_ebbinghaus$IEScongruent_vs_mild)
IES_ebbinghaus_post_ms <- describe_posterior(cor_ebbinghaus$IESmild_vs_strong)

IES_mullerlyer_post_cm <- describe_posterior(cor_mullerlyer$IEScongruent_vs_mild)
IES_mullerlyer_post_ms <- describe_posterior(cor_mullerlyer$IESmild_vs_strong)

IES_verticalhorizontal_post_cm <- describe_posterior(cor_verticalhorizontal$IEScongruent_vs_mild)
IES_verticalhorizontal_post_ms <- describe_posterior(cor_verticalhorizontal$IESmild_vs_strong)
```

>>>>

# collapsed_groups_data <- data_vs |>
#  filter(Illusion_Group %in% c("Mild", "Strong")) |>
#   group_by(Participant, Illusion_Type) |>
#   summarise(
#     Illusion_Group = "Incongruent",
#     ErrorRate = mean(ErrorRate, na.rm = TRUE),
#     IES = mean(IES, na.rm = TRUE),
#     .groups = "drop"
#   )

# data_vs <- bind_rows(data_vs, collapsed_groups_data) |>
#   arrange(Participant, Illusion_Type, factor(Illusion_Group, levels = c("Mild", "Strong", "Congruent", "Incongruent"))) |>
#   filter(Illusion_Group == c( "Congruent", "Incongruent"))

@@

```{r}
data_vs_wide <- data_vs|>
  select(Participant, Illusion_Type, Illusion_Group, ErrorRate, IES) |>
  unite("group", Illusion_Type, Illusion_Group, sep = "_") |>
  pivot_wider(names_from = group, values_from = c(ErrorRate, IES)) |>
  as.data.frame()
```

@@

**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended.

@@

  # Congruent vs Incongruent if correlation high between mild and strong

  # ERcong_vs_incon < BayesFactor::ttestBF(
  #   x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )
  # 
  # IEScong_vs_incon <- BayesFactor::ttestBF(
  #   x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$IES[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )

  return(list(
    ERcongruent_vs_mild = ERcongruent_vs_mild,
    ERmild_vs_strong = ERmild_vs_strong,
    IEScongruent_vs_mild = IEScongruent_vs_mild,
    IESmild_vs_strong = IESmild_vs_strong
    # ERcong_vs_incon = ERcong_vs_incon,
    # IEScong_vs_incon = IEScong_vs_incon
  ))
}

@@

ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
ttest_mullerlyer <- compute_ttest(data_vs, "MullerLyer")
ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
```

@@

er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
# er_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$ERcong_vs_incon)

er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
er_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
# er_mullerlyer_post_ci <- describe_posterior(ttest_mullerlyer$ERcong_vs_incon)

er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
er_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$ERmild_vs_strong)
# er_verticalhorizontal_post_ci <- describe_posterior(ttestcor_verticalhorizontal$ERcong_vs_incon)

@@

IES_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$IEScongruent_vs_mild)
IES_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$IESmild_vs_strong)
# IES_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$IEScongruent_vs_mild)
IES_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$IESmild_vs_strong)
# IES_mullerlyer_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$IEScongruent_vs_mild)
IES_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$IESmild_vs_strong)
# IES_verticalhorizontal_post_ci <- describe_posterior(ttest_verticalhorizontal$IEScong_vs_incon)
```

**Note**. Reliability analyses are conducted on all resulting indices.

#### Split-half

compute_all_split_half <- function(df, score_vars = c("ErrorRate", "IES")) {

  conditions <- c("Congruent", "Strong", "Mild")
  # conditions <- c("Congruent", "Incongruent")
  illusions <- c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")
  pairs <- list(
    c("Ebbinghaus", "MullerLyer"),
    c("VerticalHorizontal", "MullerLyer"),
    c("VerticalHorizontal", "Ebbinghaus")
  )

  # Loop through score vars and compute
  results <- list()

  for (score in score_vars) {
    cat("\n###", score, "\n")

    for (cond in conditions) {
      cat("\n##", cond, "\n")

      for (pair in pairs) {
        var1 <- paste(score, pair[1], cond, sep = "_")
        var2 <- paste(score, pair[2], cond, sep = "_")

        if (all(c(var1, var2) %in% colnames(df))) {
          cat("Pair:", var1, "vs", var2, "\n")
          res <- performance::item_split_half(df[, c(var1, var2)])
          results[[paste(score, cond, paste(pair, collapse = "_"), sep = "_")]] <- res
        } else {
          warning(paste("Missing columns:", var1, "or", var2))
        }
      }
    }
  }

  return()
}

split_half_results  <- compute_all_split_half(data_vs_wide)

#### Inter-illusion Reliability

Error Rate

con_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_Congruent")) |>
  psych::alpha()

mild_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_mild")) |>
  psych::alpha()

strong_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_strong")) |>
  psych::alpha()

# inc_rel_er <- data_vs_wide |>
#   select(matches("ErrorRate.*_Inongruent")) |>
#   psych::alpha()

con_rel_IES <- data_vs_wide |>
  select(matches("IES.*_Congruent")) |>
  psych::alpha()

mild_rel_IES <- data_vs_wide |>
  select(matches("IES.*_mild")) |>
  psych::alpha()

strong_rel_IES <- data_vs_wide |>
  select(matches("IES.*_strong")) |>
  psych::alpha()

# inc_rel_IES <- data_vs_wide |>
#   select(matches("IES.*_Inongruent")) |>
#   psych::alpha()

## PCS and Illusion Sensitivity

#### Ebbinghaus 
```{r}

cor_pcs_er_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Congruent)

er_pcs_ebb_con_post <- describe_posterior(cor_pcs_er_ebb_con)

cor_pcs_er_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Mild)

er_pcs_ebb_mild_post <- describe_posterior(cor_pcs_er_ebb_m)

cor_pcs_er_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Strong)

er_pcs_ebb_str_post <- describe_posterior(cor_pcs_er_ebb_s)

# Incongruent
# cor_pcs_er_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# er_pcs_ebb_inc_post <- describe_posterior(cor_pcs_er_ebb_inc)

#### Muller-Lyer
```{r}

cor_pcs_er_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Congruent)

er_pcs_mul_con_post <- describe_posterior(cor_pcs_er_mul_con)

cor_pcs_er_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Mild)

er_pcs_mul_mild_post <- describe_posterior(cor_pcs_er_mul_m)

cor_pcs_er_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Strong)

er_pcs_mul_str_post <- describe_posterior(cor_pcs_er_mul_s)

# Incongruent
# cor_pcs_er_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# er_pcs_mul_inc_post <- describe_posterior(cor_pcs_er_mul_inc)

#### VerticalHorizontal
```{r}

cor_pcs_er_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Congruent)

er_pcs_vh_con_post <- describe_posterior(cor_pcs_er_vh_con)

cor_pcs_er_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Mild)

er_pcs_vh_mild_post <- describe_posterior(cor_pcs_er_vh_m)

cor_pcs_er_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Strong)

er_pcs_vh_str_post <- describe_posterior(cor_pcs_er_vh_s)

# Incongruent
# cor_pcs_er_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# er_pcs_vh_inc_post <- describe_posterior(cor_pcs_er_vh_inc)
```

#### Ebbinghaus 
```{r}

cor_pcs_IES_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_pcs_ebb_con_post <- describe_posterior(cor_pcs_IES_ebb_con)

cor_pcs_IES_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_pcs_ebb_mild_post <- describe_posterior(cor_pcs_IES_ebb_m)

cor_pcs_IES_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_pcs_ebb_str_post <- describe_posterior(cor_pcs_IES_ebb_s)

# Incongruent
# cor_pcs_IES_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# IES_pcs_ebb_inc_post <- describe_posterior(cor_pcs_IES_ebb_inc)

#### Muller-Lyer
```{r}

cor_pcs_IES_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_mul_con)

cor_pcs_IES_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_mul_m)

cor_pcs_IES_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_pcs_mul_str_post <- describe_posterior(cor_pcs_IES_mul_s)

# Incongruent
# cor_pcs_IES_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# IES_pcs_mul_inc_post <- describe_posterior(cor_pcs_IES_mul_inc)
```

#### Vertical Horizontal
```{r}

cor_pcs_IES_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_vh_con)

cor_pcs_IES_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_vh_m)

cor_pcs_IES_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_pcs_vh_str_post <- describe_posterior(cor_pcs_IES_vh_s)

# Incongruent
# cor_pcs_IES_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# IES_pcs_vh_inc_post <- describe_posterior(cor_pcs_IES_vh_inc)
```

## Correlations between Illusion scores and Psychoticism

#### Ebbinghaus 
```{r}

cor_pcs_psy_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_psy_ebb_con_post <- describe_posterior(cor_pcs_psy_ebb_con)

cor_pcs_psy_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_psy_ebb_mild_post <- describe_posterior(cor_pcs_psy_ebb_m)

cor_pcs_psy_ebb_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_psy_ebb_str_post <- describe_posterior(cor_pcs_psy_ebb_s)

# Incongruent
# cor_pcs_psy_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_Ebbinghaus_Incongruent)
# 
# IES_psy_ebb_inc_post <- describe_posterior(cor_pcs_psy_ebb_inc)
```

#### Muller-Lyer 
```{r}

cor_pcs_psy_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_psy_mul_con_post <- describe_posterior(cor_pcs_psy_mul_con)

cor_pcs_psy_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_psy_mul_mild_post <- describe_posterior(cor_pcs_psy_mul_m)

cor_pcs_psy_mul_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_psy_mul_str_post <- describe_posterior(cor_pcs_psy_mul_s)

# Incongruent
# cor_pcs_psy_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_MullerLyer_Incongruent)
# 
# IES_psy_mul_inc_post <- describe_posterior(cor_pcs_psy_mul_inc)

#### Vertical-Horizontal
```{r}

cor_pcs_psy_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_psy_vh_con_post <- describe_posterior(cor_pcs_psy_vh_con)

cor_pcs_psy_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_psy_vh_mild_post <- describe_posterior(cor_pcs_psy_vh_m)

cor_pcs_psy_vh_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_psy_vh_str_post <- describe_posterior(cor_pcs_psy_vh_s)

# Incongruent
# cor_pcs_psy_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_VerticalHorizontal_Incongruent)
# 
# IES_psy_vh_inc_post <- describe_posterior(cor_pcs_psy_vh_inc)
